### 2015-02-10 paper_40_mpi.tex
working on Chapter 4

    Software development organizations struggle to measure overall performance.  
    The Master Performance Indicator (MPI) is an algorithm to provide a single 
    number score to measure the performance of a software development organization. 
    It works by statistically analyzing the past performance of the 
    organization and using that information to score an organization 
    on current performance.  MPI focuses on the following 
    elements of a software development organization: \textit{quality}, 
    \textit{availability}, \textit{satisfaction}, \textit{schedule},
    and \textit{requirements}. A separate MPI
    score is calculated for each element and 
    then aggregated together to form an overall 
    MPI score.   \textbf{MPI is not meant to be comparative between
    organizations, but to measure the amount of increase or decrease
    a single organization exhibits across elements}.  MPI is made to be easily 
    expandable to other elements if desired.

    The scores for MPI will range from -100, indicating the worst 
    performance, all the way to +100, indicating perfection.
    A score of 0 is an indication of meeting the basic expectations. 
    A negative score indicates under-performance and a 
    positive score indicates over-performance.   Here are some 
    examples. An MPI score of 35 means the organization is 
    performing 35\% better than expected.  Conversely, a score 
    of -15 means an organization is performing 15\% worse than expected.

Here are the attributes of the MPI scoring.
    \begin{itemize}
        \item The range of scores must have equal values above and below 0
        \item The minimum score must equate to the worst possible 
            performance, however that is defined
        \item Similarily, the maximum score must equate to the best possible performance.  
        \item A score of 0 must be average (or expected) performance
        \item All individual elements must have the same scoring range
    \end{itemize}

@@

\subsection{QUALITY}
    Measuring quality is a crucial part of accessing 
    software development results.  
    Poor quality means time, money, and resources are spent 
    fixing the problems. As a result, new features are not being
    created. One of the key indicators of software quality 
    is defects.  Thus, it is important
    to measure the number of defects associated with a software 
    release.  The following is the necessary data to collect
    and an algorithm to calculate
    the quality score of software being released.

    \subsubsection{QUALITY DATA}
        In order to properly score the quality of an SDO,
        certain data needs to be obtained in order to measure
        performance. Table \ref{tab:qualitydata} identifies
        the columns of data that will be used to create
        a score for the quality element of MPI.  
        Each column is classified as \textit{required} or
        \textit{optional}.  That is to allow some flexibility
        in the model for organizations that collect
        varying amounts of data.

@@

            Application ID & String (factor) & Required \\
            \myrowcolour%
            Frequency Date & Date & Required \\
            Development Effort & Integer & Required \\
            \myrowcolour%
            Testing Effort & Integer & Optional \\ 
            Defects in SIT & Integer & Optional \\ 
            \myrowcolour%
            Defects in UAT & Integer  & Optional \\ 
            Defects in PROD & Integer  & Required \\

@@

    \subsubsection{QUALITY FORMULA}
        The first step in creating a score for the quality
        element is analysis of the historical data.  The historical
        data is all quality data collected before  a given point
        in time.  Some common historical cutoffs are the current date
        or the end of the previous fiscal year.  Then a mathematical
        model to predict Defects in PROD will be produced.  In statistical
        terms, the response is \textit{Defects in PROD} and the predictors
        are: \textit{Defects in UAT}, \textit{Defects in SIT}, 
        \textit{Testing Effort}, and \textit{Development Effort}.  
        Some of the following strategies will be employed to find 
        a reasonable model.
        \begin{itemize}
            \item Removal of outliers and/or influential points
            \item Linear Regression
            \item Stepwise Regression
            \item Ridge Regression for suspected multicolinearity
        \end{itemize}

        Once a model has been found, it will be labeled as $f$ and it
        will not change.  It serves as the quality baseline for MPI. 
        All future quality scores will be dependent upon the original
        $f$.

        After the model $f$ has been determined,

\begin{displaymath}
   S_{1_i} = \left\{
     \begin{array}{lr}
       \text{where } f \geq d_i & : \left[ \frac{f - d_i}{f} \right] \\
       \text{where } d_i > f  & : \left[ \frac{f-d_i }{6\sigma} \right]
     \end{array}
   \right]
\end{displaymath}

\[
    S_{1} = \frac{\sum^n_{i=1} S_{1_i}}{n}
\]
\begin{itemize}
\item $n$ is the number of apps
\item $d_i$ is the actual defects
\item $f(h_i)$ is the function to predict defects based upon hours and other aspects of software development
\item $6\sigma$ 6 times an estimate of the standard deviation of the population using function $f$ above
\end{itemize}

There are 2 parts here.  First the system needs to be loaded with this data.  Hopefully,
a file with this data will exist going back a few months or years.  It is also important
to note that the data load must be optional if an organization does not have existing
data.  Second, every release (month/year) a new file will be upload with new information
about the most recent release.

\subsection{AVAILABILITY}
All the new requirements and great quality do not matter if the software is not available.  
The following algorithm will calculate the score of software availability.
\begin{displaymath}
   S_{3_i} = \left\{
     \begin{array}{lr}
       \text{where } A_{a_i} \leq A_{e_i} & : \left[ \frac{A_{a_i} - A_{e_i}}{A_{e_i}} \right] \\
       \text{where } A_{a_i} > A_{e_i}  & : \left[ \frac{A_{a_i} - A_{e_i} }{1-A_{e_i}} \right]
     \end{array}
   \right.
\end{displaymath} 
\[
    S_3 = \frac{\sum^n_{i=1}S_{3_i}}{n}
\]

@@

\caption{AVAILABILITY - INPUTS}
\label{tab:avail}
\end{longtable}

\subsection{SATISFACTION}
The satisfaction of the customer is also very important.  The following is an algorithm to calculate the satisfaction of software.
\[
    S_2 = \frac{\sum^n_{i=1}\left( \frac{\sum^m_{j=1}a_{ij}- \frac{min + max}{2}}{m}  \right)}{n}
\]

@@

\caption{SATISFACTION - INPUTS}
\label{tab:satisfaction}
\end{longtable}

\subsection{SCHEDULE}

        Delivery of software in a timely manner is an essential part of being a 
        successful SDO.  Being able to meet scheduled deadlines is a sign of accurate
        estimation and planning.  Drastically missing deadlines is a sign of an SDO
        with a process that needs refinement.  Studies have shown that 
        software projects exceed the estimates by an average of
        30\% \cite{Jorgensen2014}.  Thus it is important to score SDOs on accurate
        schedule adherence. Without tracking and measuring schedule adherence, it will
        not improve.

@@

    \subsubsection{Formula}

@@

    \subsubsection{Example Calculation}
    \subsubsection{Data and Calculations}
    \subsubsection{Past Performance}
    \subsubsection{Process for the Future}

    \subsubsection{ALTERNATE APPROACH}
    The best possible score should be achieved
            when meeting the estimated date exactly.  The maximum score
            should come from the best estimate.  Then given historical
            release data, it is easy to determine an 
            average $\Delta$ between the actual and the estimated.  
            Finishing a project within that $\Delta$ should result 
            in a positive score.  Outside the $\Delta$ results in
            negative scores.  For example, a project releasing
            one day early or one day late would receive the same score
            because in both cases the estimate was missed by one day.

@@

\caption{SCHEDULE - INPUTS}
\label{tab:sched}
\end{longtable}

\subsection{REQUIREMENTS}
Requirements are desired new features or enhancements to a software product.  It is important to know how many requirements
were scheduled to be completed versus how many actually got completed.  
\[
    S_5 = \frac{\sum^n_{i=1}\left( R_{a_i} - R_{e_i} \right)}{n}
\]
\begin{itemize}
\item $R_a$ actual requirements completed for application $i$
\item $R_e$ expected requirements completed for application $i$
\end{itemize}

@@

\caption{REQUIREMENT - INPUTS}
\label{tab:req}
\end{longtable}

\subsection{FINAL MPI SCORE}
One complete overall score cannot be successful without including information from various aspects of the software development organization.  Therefore, the previous results
are combined to provide one overall score.

\[
    MPI =\sum\limits^n_{i=1} w_i S_i \text{ where } \sum\limits^n_{i=1} w_i = 1
\]

@@

For more on sensitivity analysis in statistical modeling, see \cite{Saltelli2000}.

>>>>

    Software development organizations struggle to measure overall performance.  
    The Master Performance Indicator (MPI) is an algorithm to provide a single 
    number score to measure the performance of a software development organization. 
    It works by statistically analyzing the past performance of the 
    organization and using that information to score an organization 
    on current performance.  MPI focuses on the following 
    elements of a software development organization: \textit{quality}, 
    \textit{availability}, \textit{satisfaction}, \textit{schedule},
    and \textit{requirements}. A separate MPI
    score is calculated for each element and 
    then aggregated together to form an overall 
    MPI score.  A new MPI score will be calculated based
    upon the selection of a given time period (weekly, monthly, quarterly).
    \textbf{MPI is not meant to be comparative between
    organizations, but to measure the amount of increase or decrease
    a single organization exhibits across elements}.  MPI is made to be easily 
    expandable to other elements if desired.

    The scores for MPI will range from -1, indicating the worst 
    performance, all the way to +1, indicating perfection.
    A score of 0 is an indication of meeting the basic expectations. 
    A negative score indicates under-performance and a 
    positive score indicates over-performance.   Here are some 
    examples. An MPI score of 0.35 means the organization is 
    performing 35\% better than expected.  Conversely, a score 
    of -0.15 means an organization is performing 15\% worse than expected.

    Below are the attributes of the MPI scoring.
    \begin{itemize}
        \item The range of scores must have equal values above and below 0
        \item The minimum score must equate to the worst possible 
            performance, however that is defined
        \item Similarily, the maximum score must equate to the best possible performance.  
        \item A score of 0 must be average (or expected) performance
        \item All individual elements must have the same scoring range
    \end{itemize}

@@

    \subsection{ELEMENTS OF MPI}
        The MPI score consists of data collected from five
        different elements of an SDO.  
        \begin{enumerate}
            \item Quality
            \item Availability
            \item Satisfaction
            \item Schedule
            \item Requirements
        \end{enumerate}
        These five elements will be outline in the next sections.

        \subsubsection{QUALITY}
            Measuring quality is a crucial part of accessing 
            software development results.  
            Poor quality means time, money, and resources are spent 
            fixing the problems. As a result, new features are not being
            created. One of the key indicators of software quality 
            is defects.  Thus, it is important
            to measure the number of defects associated with a software 
            release.  The following is the necessary data to collect
            and an algorithm to calculate
            the quality score of software being released.

            \paragraph{QUALITY DATA}
                In order to properly score the quality of an SDO,
                certain data needs to be obtained in order to measure
                performance. Table \ref{tab:qualitydata} identifies
                the columns of data that will be used to create
                a score for the quality element of MPI.  
                Each column is classified as \textit{required} or
                \textit{optional}.  That is to allow some flexibility
                in the model for organizations that collect
                varying amounts of data.

@@

                    Application ID & String (factor) & Required \\
                    \myrowcolour%
                    Frequency Date & Date & Required \\
                    Development Effort & Integer & Required \\
                    \myrowcolour%
                    Testing Effort & Integer & Optional \\ 
                    SIT Defects & Integer & Optional \\ 
                    \myrowcolour%
                    UAT Defects & Integer  & Optional \\ 
                    PROD Defects & Integer  & Required \\

@@

            \paragraph{QUALITY FORMULA}
                The first step in creating a score for the quality
                element is analysis of the historical data.  The historical
                data is all quality data collected before  a given point
                in time.  Some common historical cutoffs are the current date
                or the end of the previous fiscal year.  Then a mathematical
                model to predict PROD Defects will be produced.  In statistical
                terms, the response is \textit{PROD Defects} and the predictors
                are: \textit{UAT Defects}, \textit{SIT Defects}, 
                \textit{Testing Effort}, and \textit{Development Effort}.  
                Some of the following strategies will be employed to find 
                a reasonable model.
                \begin{itemize}
                    \item Removal of outliers and/or influential points
                    \item Linear Regression
                    \item Stepwise Regression
                    \item Ridge Regression for suspected multicolinearity
                \end{itemize}

                Once a model has been found, it will be labeled as $f$ and it
                will not change.  The function $f$ can be the same for all 
                Application IDs or it can be different for each Application ID
                or any combination of Application IDs.   It serves as the 
                quality baseline for MPI. 
                All future quality scores will be dependent upon the original
                $f$.  Once set, the model will not change.

                After the model $f$ has been determined, it is time
                to calculate the quality score for each application ID
                within the given time period.  The quality score
                for each Application ID can be calculated as follows.

                \begin{displaymath}
                   S_{1_i} = \left\{
                     \begin{array}{lr}
                       \text{where } f_i \geq d_i & :  \frac{f_i - d_i}{f_i} \times k  \\
                       \text{where } d_i > f_i  & : \frac{f_i-d_i }{6\sigma_i} \times k
                     \end{array}
                   \right] \text{   , calculate quality score for each app $i$}
                \end{displaymath} 
                where
                \begin{itemize}
                    \item $S_{1_i}$ is the quality score for Application ID $i$
                    \item $n$ is the number of Application IDs
                    \item $d_i$ is the actual PROD defects
                    \item $f_i$ is the function to predict PROD defects for Application $i$ based upon 
                        \textit{UAT Defects}, \textit{SIT Defects}, \textit{Testing Effort}, 
                        and \textit{Development Effort}
                    \item $6\sigma$ 6 times an estimate of the standard deviation of the 
                        population using function $f_i$ above
                \end{itemize}

                \[
                    S_{1} = \sum\limits^n_{i=1} w'_i S_{1_i} \text{ where } \sum\limits^n_{i=1} w'_i = 1
                \]
                where
                \begin{itemize}
                    \item $S_1$ is the combined quality score for all Application IDs, 
                    a weighted average
                \end{itemize}

                Then $S_1$ represents the quality score for that given time period.

        \subsubsection{AVAILABILITY}
            All the new requirements and great quality do not matter if the software is not available.  
            The following algorithm will calculate the score of software availability.
            \begin{displaymath}
               S_{3_i} = \left\{
                 \begin{array}{lr}
                   \text{where } A_{a_i} \leq A_{e_i} & : \left[ \frac{A_{a_i} - A_{e_i}}{A_{e_i}} \right] \\
                   \text{where } A_{a_i} > A_{e_i}  & : \left[ \frac{A_{a_i} - A_{e_i} }{1-A_{e_i}} \right]
                 \end{array}
               \right.
            \end{displaymath} 
            \[
                S_3 = \frac{\sum^n_{i=1}S_{3_i}}{n}
            \]

@@

            \caption{AVAILABILITY DATA NEEDED FOR MPI}
            \label{tab:avail}
            \end{longtable}

        \subsubsection{SATISFACTION}
            The satisfaction of the customer is also very important.  
            The following is an algorithm to calculate the satisfaction of software.
            \[
                S_2 = \frac{\sum^n_{i=1}\left( \frac{\sum^m_{j=1}a_{ij}- \frac{min + max}{2}}{m}  \right)}{n}
            \]

@@

            \caption{SATISFACTION DATA NEEDED FOR MPI}
            \label{tab:satisfaction}
            \end{longtable}

        \subsubsection{SCHEDULE}  
            Delivery of software in a timely manner is an essential part of being a 
            successful SDO.  Being able to meet scheduled deadlines is a sign of accurate
            estimation and planning.  Drastically missing deadlines is a sign of an SDO
            with a process that needs refinement.  Studies have shown that 
            software projects exceed the estimates by an average of
            30\% \cite{Jorgensen2014}.  Thus it is important to score SDOs on accurate
            schedule adherence. Without tracking and measuring schedule adherence, it will
            not improve.

@@

            \paragraph{SCHEDULE DATA}

                \begin{longtable}{@{}l rr rr}
                    \toprule%
                     \centering%
                     {\bfseries Column Name}
                     & {\bfseries Data Type}
                     &  \\

                    \caption{SCHEDULE DATA NEEDED FOR MPI}
                    \label{tab:sched}
                \end{longtable}

            \paragraph{SCHEDULE FORMULA}

@@

            \paragraph{ALTERNATE APPROACH}
                The best possible score should be achieved
                when meeting the estimated date exactly.  The maximum score
                should come from the best estimate.  Then given historical
                release data, it is easy to determine an 
                average $\Delta$ between the actual and the estimated.  
                Finishing a project within that $\Delta$ should result 
                in a positive score.  Outside the $\Delta$ results in
                negative scores.  For example, a project releasing
                one day early or one day late would receive the same score
                because in both cases the estimate was missed by one day.

@@

        \subsubsection{REQUIREMENTS}

            \paragraph{REQUIREMENTS DATA}

                \begin{longtable}{@{}l rr rr}
                    \toprule%
                     \centering%
                     {\bfseries Column Name}
                     & {\bfseries Data Type}
                     &  \\

@@

                    \caption{REQUIREMENTS DATA NEEDED FOR MPI}
                    \label{tab:req}
                \end{longtable}

            \paragraph{REQUIREMENTS FORMULA}
                Requirements are desired new features or enhancements 
                to a software product.  It is important to know how many requirements
                were scheduled to be completed versus how many actually got completed.  
                \[
                    S_5 = \frac{\sum^n_{i=1}\left( R_{a_i} - R_{e_i} \right)}{n}
                \]
                \begin{itemize}
                    \item $R_a$ actual requirements completed for application $i$
                    \item $R_e$ expected requirements completed for application $i$
                \end{itemize}

        \subsubsection{OVERALL MPI SCORE}
            One complete overall score cannot be successful without including 
            information from various aspects of the software development 
            organization.  Therefore, the previous results
            are combined to provide one overall score.
            \[
                MPI =\sum\limits^n_{i=1} w_i S_i \text{ where } \sum\limits^n_{i=1} w_i = 1
            \]

@@

        Some simulations will be run with data from each given distribution.  Then the MPI scores
        will be analyzed.
        For more on sensitivity analysis in statistical modeling, see \cite{Saltelli2000}.
